---
title: "Using ggExtra for Exploratory Plotting (tidyverse create assignment)"
author: "Rachel Greenlee"
date: "10/24/2020"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Introduction

This vignette will take a quick peek at two useful data exploration plot types provided in the ggExtra package using a UFO sightings dataset. We aren't going to worry about style or labels, just some quick plots to explore your data well before further analysis and presenting findings to others.


### Setup
To have some fun, I picked a UFO dataset I found on Kaggle. 

```{r}
ufo <- read.csv(url("https://raw.githubusercontent.com/rachel-greenlee/data607/master/supports/ufo.csv"))
```


Let's load the needed libraries.

```{r, message=FALSE}
library(ggplot2)
library(dplyr)
```

And here is a glimpse at the data.

```{r}
glimpse(ufo)
```


We need to clean up the dataframe so we have good clean variables to plot.

```{r}
#set datetime as a datetime varirable
ufo$datetime <- as.POSIXct(ufo$datetime, format="%m/%d/%Y %H:%M", tz=Sys.timezone())

#set seconds as numeric
ufo$duration..seconds. <- as.numeric(ufo$duration..seconds.)

#convert from seconds to minutes for readability 
ufo$duration_mins <- ufo$duration..seconds. / 60

#subset for just the variables we are eniterested in
ufo <- ufo[c("datetime", "state", "country", "shape", "duration_mins")]

#make a year only variable
ufo$year <- format(as.Date(ufo$datetime, format="%m/%d/%Y %H:%M"),"%Y")
ufo$year <- as.numeric(ufo$year)

```

Take one more glimpse at the data now that it's cleaned up.

```{r}
glimpse(ufo)
```


### Plot histograms or boxplots on scatterplot axis


```{r, message=FALSE, warning=FALSE}
library(ggExtra)
```

Here we can create the base of the chart and store it in "g", and then uses the ggMarginal function from the ggExtra package to create a scatterplot that has histograms along each axis. You can also swap out type = "histogram" for "barplot". 

```{r, warning = FALSE}
#it is common to store the base of your graph in something like "g" and then add the more advanced graph features to thase base
g <- ggplot(ufo, aes(x = year, y = duration_mins, color = country)) +
  geom_count() 

ggMarginal(g, type = "histogram", fill = "transparent")
```

We see above that we have some intense outliers with regards to the duration, we can set the y axis to 1 hour (60 minutes) if we want to zoom in. Let's try that boxplot too.

```{r, warning = FALSE}
#adding ylim to greatly decrease the y axis, so we can zoom in on the majority of the data points without changing our dataframe
g <- ggplot(ufo, aes(x = year, y = duration_mins, color = country)) +
  geom_count() +
  ylim(0, 60)

ggMarginal(g, type = "boxplot", fill = "transparent")
```


### Super quick frequency plots

Another great feature in the ggExtra package is the ability to plot super quick frequency plots without having to manipulate your dataframe first.

```{r}
#when you have a lot of labels they can be hard to read on the xaxis, coord_flip will swap the axis
plotCount(table(ufo$shape)) +
  coord_flip()
```

```{r}
plotCount(table(ufo$country)) +
  coord_flip()
```

