---
title: "607 - Tidyverse Assignment"
author: "Karim Hammoud"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

# Overview
I want to analyze the sales of video games form all around the world, I collected the data from [Kaggle]("https://www.kaggle.com/gregorut/videogamesales") all numbers are per millions

# we set up the environment

```{r message=FALSE}
library(tidyverse)
```

# We pull the data

```{r}
url <- "https://raw.githubusercontent.com/akarimhammoud/CUNY-SPS/master/607-Data-Acquisition-and-Management-CUNY-SPS-Fall2020/Tidyverse/vgsales.csv"

video_games <- read_csv(url)
head(video_games)
```

# first I will rename the sales colomns 

```{r}
video_games %>% 
 rename(North_America_Sales = NA_Sales, Europe_Sales = EU_Sales, Japan_sales = JP_Sales)
```

# Arrange from the top globaly sales to the east

```{r}
video_games %>% 
 arrange(desc(Global_Sales))
```

# Group by the Genre and calculate the min, mean, median, and max.

```{r}
video_games %>% 
 group_by(Genre) %>% 
 summarise(min = min(Global_Sales), mean = mean(Global_Sales), median = median(Global_Sales), max= max(Global_Sales)) %>% 
 arrange(desc(max))
```

# Filter the top Global sales games > 30M.
```{r}
video_games_top <- video_games %>% 
 filter(Global_Sales > 30)
video_games_top
```

# Plot the top games with globaly sales.

```{r}
video_games_top %>% 
 ggplot(aes(Global_Sales, Name)) + geom_point()
```

# group global sales by year and summurize the total
```{r}
per_year <- video_games %>% 
 group_by(Year) %>% 
 summarise(Total = sum(Global_Sales)) %>% 
 filter(Year >= 2000, Year <= 2015)

per_year
```


# plot per year between 2000 - 2015

```{r}
per_year %>% 
 ggplot(aes(Year, Total)) + geom_point()
```

Any additional analysis is welcome.
Thanks,
Karim Hammoud


# Tidyverse Extend

Rachel here to add some further analysis to the great work Karim has done on one of my favorite hobbies, video games!

```{r}
#create a new dataframe using the 'subset' function to select only 'modern' PC games, because I'm a PC gamer
pc_games <- subset(video_games, Platform == "PC")
pc_games <- subset(pc_games, Year >= 2010)

#and lets remove any games with an N/A year, because that won't plot well later
pc_games <- subset(pc_games, Year != "N/A")
```


Lets see how the sales by genre have changed since 2010 at all, building a animated bubble chart using the 'gganimate' library to extend ggplot2's functionality.

```{r, message = FALSE}
# package to extend ggplot2 functionality
library(gganimate)

# package that will display the animated output when you knit to HTML
library(gifski)
```


```{r}
# built a basic ggplot barchart as you usually would
ggplot(pc_games, aes(x=Global_Sales, y=Genre, fill=Genre)) + 
  geom_bar(stat='identity') +
  theme_bw() +
        
# code specific to gganimate with "Year" as the variable it transitions on, and a label that display the years as they progress
  transition_states(Year, transition_length = 3, state_length = 10) +
  ease_aes('sine-in-out') + 
  labs(title = 'PC Game Sales by Year: {closest_state}', x = 'Global Sales', y = 'Game Genre')

```



