---
title: "Tidyverse"
author: "Orli Khaimova"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(tidyverse)
```

`fct_rev` relevels the levels of a factor in reverse order. In this case, I factored the regions and then used the function in order to put them in reverse alphabetical order. By doing so, I was able to print the regions alphabetically in the graph.

`geom_pointrange` graphs the interval for the average price of avocados for each region. I had to define a ymin and ymax as well. It is useful for drawing confidence intervals and in this case the range of prices.

```{r fig.height = 10, fig.width = 5}
avocados <- read.csv("https://raw.githubusercontent.com/okhaimova/DATA-607/master/avocado.csv")

avocados$Date <- as.Date(avocados$Date)

avocados$year <- as.character(avocados$year)

#factors the regions and then using forcats, we reverse the order to make it z-a
avocados$region <- avocados$region %>%
  as.factor() %>%
  fct_rev()

avocados %>% 
  ggplot(aes(y = AveragePrice, x = region, 
             ymin = AveragePrice-sd(AveragePrice), ymax = AveragePrice+sd(AveragePrice))) +
  geom_pointrange(aes(color = as.factor(region)), size =.01) +
  ylab("Average Price") +
  xlab("Region") +
  ggtitle("Average Price Range") +
  coord_flip() +
  theme(legend.position = "none")
```

