---
title: "Tidyverse"
author: "Orli Khaimova"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(tidyverse)
```
Overall 23/25
I think this part, "Write a vignette using more than one TidyVerse packages" is missing in your work. 
# I am a bit lost here...maybe telling your reader what you are about to do (Description). one or two lines...
# R uses factors to handle categorical variables, variables that have a fixed and known set of possible values. Factors are also helpful for reordering character vectors to improve display. The goal of the forcats package is to provide a suite of tools that solve common problems with factors, including changing the order of levels or the values. Some examples include:
https://rdrr.io/cran/forcats/f/README.md


`fct_rev` relevels the levels of a factor in reverse order. In this case, I factored the regions and then used the function in order to put them in reverse alphabetical order. By doing so, I was able to print the regions alphabetically in the graph.

`geom_pointrange` graphs the interval for the average price of avocados for each region. I had to define a ymin and ymax as well. It is useful for drawing confidence intervals and in this case the range of prices.

```{r fig.height = 10, fig.width = 5}
avocados <- read.csv("https://raw.githubusercontent.com/okhaimova/DATA-607/master/avocado.csv")

# I think it is good to add some comment so that others understand better your approach or code.
# Example here, I think you want to redefine the data type to date
avocados$Date <- as.Date(avocados$Date)
# redefine year (data type) to character...I thought ideally year is numerical or integer as data type.
# So , I can see you want to do something special with year but I cannot tell
avocados$year <- as.character(avocados$year)

#Good comment
#factors the regions and then using forcats, we reverse the order to make it z-a
avocados$region <- avocados$region %>%
  as.factor() %>%
  fct_rev()

#never done this long grouping for plot...looks good...actually reviewing your code kind of telling me about my own mistake...There is no way to see the result if I don't run ...I think there was option to include a sample result or html version...
# avocados$Date vector contains about 1000 values...a little heavy...I had that problem too.

avocados %>% 
  ggplot(aes(y = AveragePrice, x = region, 
             ymin = AveragePrice-sd(AveragePrice), ymax = AveragePrice+sd(AveragePrice))) +
  geom_pointrange(aes(color = as.factor(region)), size =.01) +
  ylab("Average Price") +
  xlab("Region") +
  ggtitle("Average Price Range") +
  coord_flip() +
  theme(legend.position = "none")
```
