---
title: "Tidyverse CREATE Assignment"
author: "Stefano Biguzzi"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Assignment
In this assignment we were asked to create a vignette to discuss one or more packages within the tidyverse library. For this assignment I chose to look at the `dplyr` library specifically the `group_by()`, `tally()`, and the `summarise()` functions.


# `tidyverse`
## Intro
We'll start the assignment with loading the `tidyverse` package. `tidyverse` encompasses 8 different packages. Some of them are:  

* `ggplot2` - for visualization
* `stringr` - for string manipulation
* `tidyr` - for data cleaning and tidying
* `dplyr` - for data manipulation
* `readr` - for reading in rectangular data like csv

## Loading
Below is the full list of the 8 packages that the `tidyverse` library includes
```{r, warning=F}
library(tidyverse)
```
Then load in the example data that was downloaded from [Kaggle](https://www.kaggle.com/andrewmvd/heart-failure-clinical-data)
```{r}
heart_failure <- read.csv(url("https://raw.githubusercontent.com/sbiguzzi/FALL2020TIDYVERSE/master/heart_failure_clinical_records_dataset.csv"))
```
```{r, echo=F}
column_list <- c("anaemia","diabetes","high_blood_pressure","smoking","DEATH_EVENT")
for(i in column_list){
  heart_failure[[i]] <- ifelse(heart_failure[[i]]==1,i,paste0("no_",i))
}
heart_failure$sex <- ifelse(heart_failure$sex==0,"Female","Male")

rm(i, column_list)
```
## The `dplyr` library
`dplyr` is the `tidyverse`'s data manipulation package. It encompasses many functions or *verbs* that can be used to solve some common data manipulation tasks. The verbs I want to focus on today are:  

* `group_by()`
* `tally()`
* `summarise()`

### `group_by()`
One of the most powerful function of dplyr library is the `group_by()`. It takes a data frame and allows you to group by one or more variables in the data frame.^[group_by() is usually used to group by a character variable]
```{r}
by_sex <- heart_failure %>% group_by(sex)
by_sex_diabetes <- heart_failure %>% group_by(sex,diabetes)
```

When you print the groupings you can see the number of groups:
```{r}
by_sex
```
Notice how it says *Groups:   sex [2]*, because there are 2 groups, Female and Male. When you print the grouping by sex and diabetes you get 4 groups:
```{r}
by_sex_diabetes
```
Notice how it shows *Groups:   sex, diabetes [4]* since its doing female-diabetes, male-diabetes, female-no_diabetes, and male-no_diabetes.

### `tally()`
Tally works well with the `group_by()` verb as it allows you to count the number of rows within each group. If we go back and use the same grouping *sex* and *sex_diabetes* we see the following counts:
```{r}
heart_failure %>% group_by(sex) %>% tally()
```
We can see the 2 groups that `group_by()` create *Female* and *Male* and see the number of rows for each group.
```{r}
heart_failure %>% group_by(sex,diabetes) %>% tally()
```
Here we see the 4 groups that `group_by()` created and the number of rows for each of those.

### `summarise()`

The `summarise()` function allows you manipulate the data within each group. For example if you want to get the mean platelets by sex you would use `summarise()` like this:

```{r, message=F}
heart_failure %>%
  group_by(sex) %>%
  summarise(mean_platelets = mean(platelets))
```
It's possible to also get multiple statistics using the `summarise()` function. For example below I am grabbing the mean, median, min, max, and standard deviation for platelets grouped by the sex of the patient:
```{r, message=F}
heart_failure %>%
  group_by(sex) %>%
  summarise(
    min_platelets = min(platelets),
    max_platelets = max(platelets),
    mean_platelets = mean(platelets),
    median_platelets = median(platelets),
    sd_platelets = sd(platelets)
  )
```