---
title: "TidyVerse EXTEND"
author: "Stefano Biguzzi"
date: "10/23/2020"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part I: Arushi Arora - CREATE

The following code was written by Arushi Arora on 10/23 for her TidyVerse CREATE assignment.

## Introdction
The following example from Five Thirty Eight https://fivethirtyeight.com/features/how-americans-like-their-steak/

The core tidyverse package includes "readr" and "dplyr"
"readr" provides a fast and friendly way to read rectangular data
"dplyr" provides a grammar of data manipulation, providing a consistent set of verbs that solve the most common data manipulation 

#### Importing libraries
```{r warning=FALSE}
if(!require('tidyverse'))install.packages('tidyverse')
library (tidyverse)
library(dplyr, warn.conflicts = FALSE) # for cleaning the data
options(dplyr.summarise.inform = FALSE)
```


```{r warning=FALSE}
urlfile="https://raw.githubusercontent.com/fivethirtyeight/data/master/steak-survey/steak-risk-survey.csv"

steakdata<- readr::read_csv(url(urlfile))

head(steakdata)
```

## Rename columns

```{r warning=FALSE}
steakdata1 = dplyr::rename(steakdata, 
"lottery" = "Consider the following hypothetical situations: <br>In Lottery A, you have a 50% chance of success, with a payout of $100. <br>In Lottery B, you have a 90% chance of success, with a payout of $20. <br><br>Assuming you have $10 to bet, would you play Lottery A or Lottery B?", 
"smoke_cigs" = "Do you ever smoke cigarettes?" ,
"drink_alcohol" = "Do you ever drink alcohol?", 
"gamble" = "Do you ever gamble?",
"skydiving" = "Have you ever been skydiving?",
"overspeeding" = "Do you ever drive above the speed limit?",
"cheat_patner" = "Have you ever cheated on your significant other?",
"eat_steak" = "Do you eat steak?",
"steak_prep" = "How do you like your steak prepared?",
"hh_income" = "Household Income",
"location" = "Location (Census Region)")
head(steakdata1)
```

## Remove first row

```{r warning=FALSE}
steakdata2 <- steakdata1[-c(1), ]

head(steakdata2)
```

## Mutate Char variables to Factor

```{r warning=FALSE}
steakdata3 <- steakdata2 %>% as_tibble() 

steakdata4 <- steakdata3 %>%
  mutate(lottery = as.factor(lottery)) %>%
  mutate(smoke_cigs = as.factor(smoke_cigs)) %>%
  mutate(drink_alcohol = as.factor(drink_alcohol)) %>%
  mutate(gamble = as.factor(gamble)) %>%
  mutate(skydiving = as.factor(skydiving)) %>%
  mutate(overspeeding = as.factor(overspeeding)) %>%
  mutate(cheat_patner = as.factor(cheat_patner)) %>%
  mutate(eat_steak = as.factor(eat_steak)) %>%
  mutate(steak_prep = as.factor(steak_prep)) %>%
  mutate(Gender = as.factor(Gender)) %>%
  mutate(Age = as.factor(Age)) %>%
  mutate(hh_income = as.factor(hh_income)) %>%
  mutate(Education = as.factor(Education)) %>%
  mutate(location = as.factor(location))

  head(steakdata4)
```

# Part II: Stefano Biguzzi - EXTEND

### Introduction
Your task here is to **Extend an Existing Example**.  Using one of your classmate’s examples (as created above), extend his or her example with additional annotated code.

### `ggplot()`
Once you use `readr` and `dplyr` you can use another TidyVerse package, `ggplot`, to create visualizations with your data. The following data lends itself to bar graphs.

We start out with a simple bar graph using `ggplot()` and `geom_bar()`. We set the x = to steak_prep and the y = n. In the `geom_bar()` options we add `stat = identity` to tell R that we are providing the y values for the graph.
```{r, fig.show='asis', fig.width = 6}
filter(steakdata4,!(is.na(steakdata4$steak_prep))) %>%
  count(steak_prep) %>%
  #Initiate ggplot
  ggplot(aes(x=steak_prep,y=n)) +
  #tell ggplot you want a bar graph
  geom_bar(stat = 'identity') +
  #set the theme
  theme_bw()
```

This graph does not look particularly appealing. Luckily, `ggplot` allows for a lot of customization to spice up your plots. You can color your groups by adding the `fill=` in the aes function of `ggplot`. In our example we add `fill=steak_prep` to give each bar a separate color. We are also reordering the x-axis so the cook level is ordered in descending order.
```{r, fig.show='asis', fig.width = 6}
filter(steakdata4,!(is.na(steakdata4$steak_prep))) %>%
  count(steak_prep) %>%
  #Initiate ggplot, add color and reorder 
  ggplot(aes(x=reorder(steak_prep,-n),y=n, fill=steak_prep, na.rm = TRUE)) +
  #tell ggplot you want a bar graph
  geom_bar(stat="identity", na.rm = TRUE) +
  #set the theme
  theme_bw()
```

`ggplot` allows you to add labels for each axis so a messy column name does not have to show up on a professional graph. In our bar example, we can add the `labs()` function for the title and the axis labels. We also added `show.legend = FALSE` since the x-axis is already acting as a sort of label.
```{r, fig.show='asis', fig.width = 6}
filter(steakdata4,!(is.na(steakdata4$steak_prep))) %>%
  count(steak_prep) %>%
  #Initiate ggplot, add color and reorder 
  ggplot(aes(x=reorder(steak_prep,-n),y=n)) +
  #tell ggplot you want a bar graph
  geom_bar(aes(fill=steak_prep), stat="identity", show.legend = FALSE) +
  #add titles
  labs(title = "Count of steak prep", x = "Steak Prep", y = "Counts") +
  #set the theme
  theme_bw()
```

The theme options in `ggplot` can help you further customize a plot. Let's say we like the way the black and white theme works, but we want to remove the x-axis grid. In our bar plot example below we can add an additional theme element that does just that.
```{r, fig.show='asis', fig.width = 6}
filter(steakdata4,!(is.na(steakdata4$steak_prep))) %>%
  count(steak_prep) %>%
  #Initiate ggplot, add color and reorder 
  ggplot(aes(x=reorder(steak_prep,-n),y=n)) +
  #tell ggplot you want a bar graph
  geom_bar(aes(fill=steak_prep), stat="identity", show.legend = FALSE) +
  #add titles
  labs(title = "Count of steak prep", x = "Steak Prep", y = "Counts") +
  #set the theme
  theme_bw() +
  #remove grid background
  theme(panel.grid.major.x = element_blank())

```

### Conclusion
`ggplot` is a very powerful tool for visualizations. In this example I only showed one way of visualizing data, bar plots. However, `ggplot` has a variety of layers that can be created. For a full list of available plots, or a full list of available functions to customize your plots even further you may want to read up on some [documentation](https://ggplot2.tidyverse.org/reference/).
