---
title: "Using Tidyverse readr"
author: "George Cruz"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

**readr** is a package of tidyverse that provides the ability to import data from common data files like: csv, tsv and fixed-wide-format.  It is flexible enough to handle most of these files but it will also elegantly fail if something is not right. 


## Using **readr**

### Installation

The [official web site](https://readr.tidyverse.org/) tells us: 
```{r eval=FALSE}
# The easiest way to get readr is to install the whole tidyverse:
install.packages("tidyverse")

# Alternatively, install just readr:
install.packages("readr")

```

### Using the library

Since **readr** is part of tidyverse, loading it is as simple as loading the tidyverse:

```{r}
library(tidyverse)
```

### Reading Data

The functions to read files share the same arguments and follow the same structure: 

`read_*(file, col_names=TRUE, col_types=NULL, locale=default_locale(), na=c("","NA"),quoted_na=TRUE,comment="",trim_ws=TRUE,skip=0,n_max=lnf,guess_max=min(1000,n_max),progress=interactive() )`

The only required argument for the function is the filename. Let's read the argument description from an actual csv.

```{r warning=FALSE}
arguments_list <- read_csv("../data/arguments_for_read_csv.csv")
knitr::kable(arguments_list)
```


### An Example

To show another example, let's read data from the Election Forecasts csv files listed at [FiveThirtyEight](https://data.fivethirtyeight.com/).  More specifically, we will read the file: `electoral_college_vs_popvote.csv`

```{r}
electoral_data <- read_csv("../data/electoral_college_vs_popvote.csv", 
                           col_types = cols(
                             cycle = col_integer(),
                             branch = col_character(),
                             model = col_character(),
                             modeldate = col_character(),
                             candidate_inc = col_character(),
                             candidate_chal = col_character(),
                             candidate_3rd = col_character(),
                             lower_bin_text = col_character(),
                             upper_bin_text = col_character(),
                             ecwin_inc = col_double(),
                             ecwin_chal = col_double(),
                             ecwin_3rd = col_double(),
                             ecwin_nomajority = col_double(),
                             total_ev_inc = col_double(),
                             ev_inc_hi = col_double(),
                             ev_inc_lo = col_double(),
                             total_ev_chal = col_double(),
                             ev_chal_hi = col_double(),
                             ev_chal_lo = col_double(),
                             count = col_double(),
                             timestamp = col_character(),
                             simulations = col_integer()
                           ))
```

Specifying the column types reduces the warnings that could have been caused by missed conversions. 

```{r}
knitr::kable(head(electoral_data))
```

This table lists the probable outcomes of the presidential election based on 4000 simulations. 


### Let's see the top 10 probable results by using arrange

```{r}
sorted_by_probabilities <- electoral_data %>%
  arrange(desc(count))

knitr::kable(head(sorted_by_probabilities,10))
```


